---
title: "Mouse_DESeq2"
author: "Matthew Rada"
date: "2024-08-23"
output: html_document
---

Lets load the required libraries for this analysis
```{r}

library("DESeq2")
library("dplyr")
library("apeglm")
```

Lets load in our data
```{r}
countData <- read.csv("GLDS-102_rna_seq_Unnormalized_Counts.csv")

colData <- read.csv("new_pheno_mouse3.csv", sep = ",", row.names = 1)
```


Now we need to add leveling factors to the colData
```{r}
colData$group <- factor(colData$group, levels = c("0", "1"))
```

Now lets make sure we have the same number of individuals as well as groups
```{R}
all(rownames(colData)%in% colnames(countData)) # its false
```

We need to round up the counts, because DESeq2 only allows integers as an input, and not fractional numbers. This depends on the method of alignment that was used upstream.
```{r, results='hide'}
countData %>%
  mutate_if(is.numeric, ceiling)

countData[2:13] <- sapply(countData[,2:13], as.integer)

row.names(countData) <- countData[,1]

countData <- countData[2:13]

rownames(colData) == colnames(countData) # its false since ur using self made csv.. should be true tho lol if real data

```


```{r}

dds <- DESeqDataSetFromMatrix(countData = countData, colData = colData, design = ~group)

dds <- dds[rowSums(counts(dds)>2) >= 4]

dds <- DESeq(dds)

res <- results(dds)
```
Could not finished tutorial due to missing data...

